---
title: "p8105_hw1_qh2284"
author: "Qin Huang"
date: "2023-09-15"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
```

# Problem 1

```{r}
library(moderndive)
data("early_january_weather")
```

```{r}
# To get a basic understanding of the dataset
?early_january_weather
head(early_january_weather)

```

-   **Discuss the variables in this dataset, including names / values of important variables**

    There are total 15 variables.

    1.  origin: it shows the weather station. (values only contain EWR)
    2.  year, month, day, hour: time of recording. (values from 2013.01.01 0 hour t0 2013.01.15 23 hour)
    3.  temp, dewp: temperature and dewpoint in F. (values varies)
    4.  humid: relative humidity. (values varies)
    5.  wind_dir, wind_speed, wind_gust: wind direction in degrees, wind speed in mph, and gust speed in mph (some missing values in wind_gust). (values varies)
    6.  precip: precipitation in inches. (values varies)
    7.  pressure: Sea level pressure in millibars. (values varies, some missing values)
    8.  visib: visibility in miles. (values varies)
    9.  time_hour: date and hour of the recording as a POSIXct date.

------------------------------------------------------------------------

```{r}
#calculate the size of the dataset

cols = early_january_weather %>%
  ncol()

rows = nrow(early_january_weather)

cols
rows
```

-   **The size of the dataset:** `early_january_weather` dataset has 358 rows and 15 cols (represents 15 variables and 358 observations)

------------------------------------------------------------------------

```{r}
# calculate the mean temperature
mean_temp = mean(early_january_weather$temp, na.rm = T) #get rid of the missing values if any

mean_temp
```

-   **The mean temperature:** `early_january_weather` dataset has a mean temperature of 39.58212 F.

------------------------------------------------------------------------

```{r}
# Make a scatterplot of temp (y) vs time_hour (x)
plot_p1 = early_january_weather %>%
  ggplot(aes(x = time_hour, y = temp, color = humid))  +
  geom_point() +
  labs(x = 'Date and hour', y = 'Temperature (F)', title = 'Scatter plot of temperature vs time_hour')

plot_p1

ggsave("Scatter plot.png", plot = plot_p1)
```

-   **The patterns that are apparent in this plot**

    The trend of temperature and humidity changes in every day is roughly the same, during the day the temperature is high and the humidity is high, and in the evening and night the temperature is low and the humidity is low.

# Problem 2

```{r}
# create a random sample of size 10 from a standard Normal distribution

set.seed(210) # set a seed to make the random results same

p2_random = rnorm(10, mean = 0, sd = 1)

p2_random
```

```{r}
# Create a logical vector indicating whether elements are greater than 0

p2_logic = p2_random > 0

p2_logic
```

```{r}
# Create a character vector of length 10

p2_char = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J")

p2_char
```

```{r}
# Create a factor vector of length 10 with 3 different levels

p2_factor = factor(rep(c("Poor", "Normal", "Good"), length.out = 10))

p2_factor
```

```{r}
# create a dataframe

p2_data = data.frame(
  RandomSample = p2_random,
  IsGT0 = p2_logic,
  Chars = p2_char,
  Factors = p2_factor
)

p2_data
```

```{r}
# calculate the means within the dataframe
random_mean = mean(pull(p2_data, RandomSample), na.rm = T) # get rid of the missing values

random_mean

logic_mean = mean(pull(p2_data, IsGT0))

logic_mean


```

```{r}

char_mean = mean(pull(p2_data, Chars))

char_mean
```

```{r}
factor_mean = mean(pull(p2_data, Factors))

factor_mean
```

Above all, we could observe that we could calculate the means for the random samples and the logical vector, but we could not calculate the means for character vector and factor vector.

```{r}
# convert logical, character, and factor variables to numeric

num_logic = as.numeric(p2_data$IsGT0)
```

```{r}
num_char = as.numeric(p2_data$Chars)
```

```{r}
num_factor = as.numeric(p2_data$Factors)
```

Above all, we could observe that we could apply `as.numeric` function to logical values and factor vectors, but could not apply to character vectors.

The reasons are as below:

1.  **Logical Variable**: Applying **`as.numeric()`** to a logical variable will convert **`TRUE`** to 1 and **`FALSE`** to 0, since it's a binary coding of logical values.

2.  **Character Variable:** There will be two situation when applying **`as.numeric()`** to a character variable. If character variables are valid numeric representations (e.g., "1", "2", "3"), they could be converted to numeric. If the character values are not numeric representations (e.g., "A", "B", "C"), they will be converted to **`NA`** (missing values). In our example, since the character values are "A", "B", ..., "J", they could NOT be converted to numeric.

3.  **Factor Variable**: There will be two situation when applying **`as.numeric()`** to a factor vector. If the factors are ordinal, they could be convert to numeric vectors where each level is replaced with its corresponding integer code. If the factors are nominal, it may not be meaningful (but it can also convert to numeric). In this example, since "poor", "normal", "good" has natural order, it's meaningful with the numeric conversion.

Since R could only calculate the means of numeric values. Except logical variable would automatically convert to numeric, factor vector and char vector could not be calculated a mean. Apply `as.numeric` function helps to calculate the mean values for different variable. But as mentioned above, some conversion is impossible and some conversion is meaningless, we should take these into consideration when we do the conversion.
